---
output:
  html_document: default
---

# Data wrangling exercise 
# Exercise number one

# set up the environment
```{r}
library(tidyr)
library(dplyr)
library(readxl)
```
## 1.  Read the document from downloads to R
```{r}
library(readxl)
refine<-read_excel("/Users/paulcarson/Downloads/refine.xlsx")

```

## 2. clean up company names
###change all of the column to lower case
```{r}
refine$company<-tolower(refine$company)
```

### check for a pattern to correct spelling mistakes.  following looks for the start of each company name.  Phillips starts with an f for one entry, so final line looks for the ps at the end to substitue
```{r}
refine$company<- sub(pattern = "^p.*", replacement = "phillips", x= refine$company)
refine$company<- sub(pattern = "^a.*", replacement = "akzo", x= refine$company)
refine$company<- sub(pattern = "^v.*", replacement = "von houten", x= refine$company)
refine$company<- sub(pattern = "^u.*", replacement = "unilever", x= refine$company)
refine$company<- sub(pattern = ".*\\ps$", replacement = "phillips", x= refine$company)
```


## 3.  separate out product code and number into different columns
```{r}
refine <- separate (refine, "Product code / number", c("product_code", "product_number"), sep = "-")
```


## 4.  add product categories
###establish a look up table based on the key from the table
###establish a column, product_category with using the look up table with the coded column

```{r}
lut<-c("p"="Smartphone","v"="TV","x"="Laptop","q"="Tablet")
refine$product_category <- lut[refine$product_code]
```

## 5.  put together columns for a full address for geo coding
```{r}
refine <- refine %>% mutate(full_address = paste(address,city,country, sep =","))
```

## 6. create dummy variables for company and product category
### four binary columns for company: company_phillips, company_akzo,company_van_houten,company_unilver
```{r}
refine<-mutate(refine,company_phillips=ifelse(company=="phillips",1,0))
refine<-mutate(refine,company_akzo=ifelse(company=="akzo",1,0))
refine<-mutate(refine,company_van_houten=ifelse(company=="von houten",1,0))
refine<-mutate(refine,company_unilver=ifelse(company=="unilever",1,0))
```


###four binary columns for product category: product_smartphone, product_tv,product_laptop,product_tablet
##save final as a csv file: refine_clean up
```{r}
refine<- mutate(refine,product_smartphone=ifelse(product_category=="smartphone",1,0))
refine<- mutate(refine,product_tv=ifelse(product_category=="tv",1,0))
refine<- mutate(refine,product_laptop=ifelse(product_category=="laptop",1,0))
refine<- mutate(refine,product_tablet=ifelse(product_category=="tablet",1,0))
write.csv(refine,"refine_clean.csv")
```




